Climate predictor selection and variable reduction methods influence MaxEnt model performance and predictions of climatic suitability

### 2023-06-05 manuscript/manuscript.qmd
Add spatial autocorrelation text in methods

---
title: Climate predictor selection and variable reduction methods influence MaxEnt model performance and predictions of climatic suitability 
author:
  - name: Clarke van Steenderen
    email: vsteenderen@gmail.com
    affiliations: 
        - id: some-tech
          name: Center for Biological Control, Department of Zoology and Entomology
          department: Rhodes University
          address: 
          city: Makhanda
          state: South Africa
          postal-code: 6140
    attributes:
        corresponding: false
  - name: Guy F. Sutton
    email: g.sutton@ru.ac.za
    affiliations:
        - id: another-u
          name: Center for Biological Control, Department of Zoology and Entomology
          department: Rhodes University
          address: 
          city: Makhanda
          state: South Africa
          postal-code: 6140
    attributes:
        corresponding: true
abstract: |
  This is the abstract. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Vestibulum augue turpis, dictum non malesuada a, volutpat eget velit. Nam placerat turpis purus, eu tristique ex tincidunt et. Mauris sed augue eget turpis ultrices tincidunt. Sed et mi in leo porta egestas. Aliquam non laoreet velit. Nunc quis ex vitae eros aliquet auctor nec ac libero. Duis laoreet sapien eu mi luctus, in bibendum leo molestie. Sed hendrerit diam diam, ac dapibus nisl volutpat vitae. Aliquam bibendum varius libero, eu efficitur justo rutrum at. Sed at tempus elit.
keywords: 
  - keyword1
  - keyword2
date: last-modified
bibliography: bibliography.bib
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      name: Ecological Modelling
      formatting: preprint
      model: 3p
      layout: twocolumn
      cite-style: authoryear
      highlights:
        - Highlight 1
        - Highlight 2 
        - Highlight 3
---

@@

@@

Climate data were obtained by downloading the standard set of 19 bioclimatic variables from the WorldClim ver. 2.1 database (Fick and Hijmans, 2017) (data available at: www.worldclim.org/download. html). This dataset is representative of annual and seasonal means and variation of temperature and precipitation metrics averaged over the 1950–2000 time period (current climate) at a 2.5 min resolution. These variables have been shown to effectively model the climatic suitability for non-native insects (e.g., Trethowan et al., 2011).

@@

MaxEnt (ver. 3.4.3) was implemented in the ‘ dismo’ R package (Hijmans et al., 2021 ).

>>>>

---
title: Climate predictor selection and variable reduction methods influence MaxEnt model performance and predictions of climatic suitability 
author:
  - name: Clarke van Steenderen
    email: vsteenderen@gmail.com
    affiliations: 
        - id: some-tech
          name: Center for Biological Control, Department of Zoology and Entomology
          department: Rhodes University
          address: 
          city: Makhanda
          state: South Africa
          postal-code: 6140
    attributes:
        corresponding: false
  - name: Guy F. Sutton
    email: g.sutton@ru.ac.za
    affiliations:
        - id: another-u
          name: Center for Biological Control, Department of Zoology and Entomology
          department: Rhodes University
          address: 
          city: Makhanda
          state: South Africa
          postal-code: 6140
    attributes:
        corresponding: true
abstract: |
  This is the abstract. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Vestibulum augue turpis, dictum non malesuada a, volutpat eget velit. Nam placerat turpis purus, eu tristique ex tincidunt et. Mauris sed augue eget turpis ultrices tincidunt. Sed et mi in leo porta egestas. Aliquam non laoreet velit. Nunc quis ex vitae eros aliquet auctor nec ac libero. Duis laoreet sapien eu mi luctus, in bibendum leo molestie. Sed hendrerit diam diam, ac dapibus nisl volutpat vitae. Aliquam bibendum varius libero, eu efficitur justo rutrum at. Sed at tempus elit.
keywords: 
  - Species distribution model
  - Variable selection
  - Ecological model
  - WORLDCLIM
date: last-modified
bibliography: bibliography.bib
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      name: Ecological Modelling
      formatting: preprint
      model: 3p
      layout: twocolumn
      cite-style: authoryear
      highlights:
        - Highlight 1
        - Highlight 2 
        - Highlight 3
---

@@

- Clarke to add information on where we sourced GPS records - link to key papers, maybe a supplementary table (if required).
- How many records do we have?

Spatial autocorrelation is an important factor that may affect model outputs. Filtering of species occurrence data may limit the inherent biases in the data and improve model quality (Veloz, 2009). To avoid pseudo-replication, only one occurrencenrecord per 2.5 min grid cell was used for model calibration. Species occurrence datasets were thinned using the ‘*spThin*’ package [@Aiello-Lammens2015], and spatial autocorrelation analyses were performed using the ‘*ecospat*’ package [@DiCola2017]. Eighty-eight occurrence records were retained to calibrate native-range models. Spatial thinning was also performed for the invaded range occurrences, however, there was no evidence for spatial autocorrelation. As such, no occurrence records were thinned for the invaded-range dataset.

@@

Climate data were obtained by downloading the standard set of 19 bioclimatic variables from the WorldClim ver. 2.1 database [@Fick2017] (data available at: www.worldclim.org/download.html). This dataset is representative of annual and seasonal means and variation of temperature and precipitation metrics averaged over the 1950–2000 time period (current climate) at a 2.5 minute resolution. These variables have been shown to effectively model the climatic suitability for non-native insects (e.g., @Sutton2022).

@@

MaxEnt (ver. 3.4.3) was implemented in the ‘*dismo*’ R package (@Hijmans2017).
